Model Adaptivity for Goal-Oriented Inference using Adjoints

### 2017-06-29 paper/submitted/round2_png/mathematical_formulation_v4_cmame_vikram_HLedit.tex
Added remark on when we actually get a third order estimate after Proposition 2.

%
\begin{proposition}
\label{thm:error_estimate_dual}
Let the hypotheses of Proposition~\ref{thm:error_estimate} hold. Let the variational form
%
\begin{equation}
\mathscr{R}(\Psi_{HF})(\Phi)=0,\quad\forall\Phi\in(Q_{HF}\times U_{HF}\times U_{HF})^2
\label{eq:supadjsys_prop}
\end{equation}
%
represent the equation $\mathcal{M}'_{HF,\Psi}(\Psi_{HF})(\Phi)=0$ in residual form, where $\Phi$ is a test function. Define an adjoint problem
%
\begin{equation}
\mathscr{R}'(\Psi_{LF},\Phi)(\Lambda)=\mathcal{Q}(\Phi),\quad\forall\Phi\in(Q_{HF}\times U_{HF}\times U_{HF})^2
\label{eq:superAdjEq}
\end{equation}
for the supplementary adjoint $\Lambda$, where $\mathscr{R}$ is defined in \Cref{eq:supadjsys_prop}, and where
%Let $\Lambda$ be the solution of the dual problem, \red{$\mathcal{M}'$ has just one '?}
%
%\begin{equation}
%\mathcal{M}_{HF}'(\Psi_{LF};\Phi, \Lambda)=\mathcal{Q}(\Phi),\quad\forall\Phi\in(Q_{HF}\times U_{HF}\times U_{HF})^2,
%\end{equation}
%
where $\mathcal{Q}:(U_{HF} \times U_{HF} \times Q_{HF})^2 \to \R$ is a linear functional defined by,
%
\begin{equation}
\mathcal{Q}(\Phi)=\mathcal{M}'_{HF}(\Psi_{LF})(\Phi).
\label{eq:supadjout}
\end{equation}
%
Then we have,
%
\begin{multline}
\label{eq:finErrExp}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\-\frac{1}{2}\mathcal{M}'_{HF}(\Psi_{LF})(\Lambda)+\mathcal M_{HF}(\Psi_{LF})-\mathcal M_{LF}(\Psi_{LF}) + \mathcal{R}_{\mathscr{R}} + \mathcal{R}
\end{multline}
%
where $\mathcal{R}_{\mathscr{R}}$ is a second-order term in the error $e=\Psi_{HF}-\Psi_{LF}$, and $\mathcal{R}$ is as defined in \Cref{thm:error_estimate}. %\red{Now that we've explicitly pointed out the second-order error term, can we really say that we derived "an adjoint-based third-order estimate" (in abstract and intro)?}
%
\end{proposition}
%
\begin{proof}
%
The proof simply follows from the classic dual-based a posteriori error estimation strategy. Note that, as a stationary point, $\Psi_{HF}$ satisfies,
%
\begin{equation}
\mathscr{R}(\Psi_{HF})(\Phi)=0,\quad\forall\Phi\in(Q_{HF}\times U_{HF}\times U_{HF})^2
\label{eq:supadjsys}
\end{equation}
%
which is represents the equation $\mathcal{M}'_{HF}(\Psi_{HF})(\Phi)=0$ in residual form, where $\Phi$ is a test function. Define the adjoint problem
%
\begin{equation}
\mathscr{R}'(\Psi_{LF},\Phi)(\Lambda)=\mathcal{Q}(\Phi),\quad\forall\Phi\in(Q_{HF}\times U_{HF}\times U_{HF})^2
\label{eq:superAdjEq}
\end{equation}
%
for the supplementary adjoint $\Lambda$. The error in the linear output $\mathcal{Q}$ defined in \Cref{eq:supadjout} can then be expressed as a dual-weighted residual,
%
\begin{equation}
\label{eq:adjOutErr}
\mathcal{Q}(\Psi_{HF}-\Psi_{LF})=-\mathscr{R}(\Psi_{LF})(\Lambda) + R_{\mathscr{R}},
\end{equation}
%
where $R_{\mathscr{R}}$ is a second order remainder term due to the linearization of $\mathscr{R}$ about $\Psi_{LF}$ in \Cref{eq:superAdjEq}. This implies the required result. %\red{What is the classical notation for adjoint problems? Is it okay that we're breaking from the previous notation where a second pair of parentheses denotes things that we are linear in?}
%
\end{proof}
%
Other approaches to estimating the term $\frac{1}{2}\mathcal{M}'_{HF}(\Psi_{LF})(\Psi_{HF}-\Psi_{LF})$ can become feasible in certain scenarios; for example, if the high- and low-fidelity models only differ in the computational grid used, the difference $\Psi_{HF}-\Psi_{LF}$ can be approximated using interpolation or patch recovery methods~\cite{BecVex05}.

>>>>

%
\begin{proposition}
\label{thm:error_estimate_dual}
Let the hypotheses of Proposition~\ref{thm:error_estimate} hold. Let the variational form
%
\begin{equation}
\mathscr{R}(\Psi_{HF})(\Phi)=0,\quad\forall\Phi\in(Q_{HF}\times U_{HF}\times U_{HF})^2
\label{eq:supadjsys_prop}
\end{equation}
%
represent the equation $\mathcal{M}'_{HF,\Psi}(\Psi_{HF})(\Phi)=0$ in residual form, where $\Phi$ is a test function. Define an adjoint problem
%
\begin{equation}
\mathscr{R}'(\Psi_{LF},\Phi)(\Lambda)=\mathcal{Q}(\Phi),\quad\forall\Phi\in(Q_{HF}\times U_{HF}\times U_{HF})^2
\label{eq:superAdjEq}
\end{equation}
for the supplementary adjoint $\Lambda$, where $\mathscr{R}$ is defined in \Cref{eq:supadjsys_prop}, and where
%Let $\Lambda$ be the solution of the dual problem, \red{$\mathcal{M}'$ has just one '?}
%
%\begin{equation}
%\mathcal{M}_{HF}'(\Psi_{LF};\Phi, \Lambda)=\mathcal{Q}(\Phi),\quad\forall\Phi\in(Q_{HF}\times U_{HF}\times U_{HF})^2,
%\end{equation}
%
where $\mathcal{Q}:(U_{HF} \times U_{HF} \times Q_{HF})^2 \to \R$ is a linear functional defined by,
%
\begin{equation}
\mathcal{Q}(\Phi)=\mathcal{M}'_{HF}(\Psi_{LF})(\Phi).
\label{eq:supadjout}
\end{equation}
%
Then we have,
%
\begin{multline}
\label{eq:finErrExp}
I(q_{HF},u_{HF})-I(q_{LF},u_{LF})=\\-\frac{1}{2}\mathcal{M}'_{HF}(\Psi_{LF})(\Lambda)+\mathcal M_{HF}(\Psi_{LF})-\mathcal M_{LF}(\Psi_{LF}) + \mathcal{R}_{\mathscr{R}} + \mathcal{R}
\end{multline}
%
where $\mathcal{R}_{\mathscr{R}}$ is a second-order term in the error $e=\Psi_{HF}-\Psi_{LF}$, and $\mathcal{R}$ is as defined in \Cref{thm:error_estimate}. %\red{Now that we've explicitly pointed out the second-order error term, can we really say that we derived "an adjoint-based third-order estimate" (in abstract and intro)?}
%
\end{proposition}
%
\begin{proof}
%
The proof simply follows from the classic dual-based a posteriori error estimation strategy. Note that, as a stationary point, $\Psi_{HF}$ satisfies,
%
\begin{equation}
\mathscr{R}(\Psi_{HF})(\Phi)=0,\quad\forall\Phi\in(Q_{HF}\times U_{HF}\times U_{HF})^2
\label{eq:supadjsys}
\end{equation}
%
which is represents the equation $\mathcal{M}'_{HF}(\Psi_{HF})(\Phi)=0$ in residual form, where $\Phi$ is a test function. Define the adjoint problem
%
\begin{equation}
\mathscr{R}'(\Psi_{LF},\Phi)(\Lambda)=\mathcal{Q}(\Phi),\quad\forall\Phi\in(Q_{HF}\times U_{HF}\times U_{HF})^2
\label{eq:superAdjEq}
\end{equation}
%
for the supplementary adjoint $\Lambda$. The error in the linear output $\mathcal{Q}$ defined in \Cref{eq:supadjout} can then be expressed as a dual-weighted residual,
%
\begin{equation}
\label{eq:adjOutErr}
\mathcal{Q}(\Psi_{HF}-\Psi_{LF})=-\mathscr{R}(\Psi_{LF})(\Lambda) + R_{\mathscr{R}},
\end{equation}
%
where $R_{\mathscr{R}}$ is a second order remainder term due to the linearization of $\mathscr{R}$ about $\Psi_{LF}$ in \Cref{eq:superAdjEq}. This implies the required result. %\red{What is the classical notation for adjoint problems? Is it okay that we're breaking from the previous notation where a second pair of parentheses denotes things that we are linear in?}
%
\end{proof}
%
If the state equations of the high fidelity model are linear, the derivative of the Lagrangian in~\Cref{eq:InvsOpt_auglag} is linear. In this case, the $R_{\mathscr{R}}$ term in~\Cref{eq:finErrExp} disappears, even if the QoI functional $I$ is nonlinear, and we get a fully third order error estimate.

Other approaches to estimating the term $\frac{1}{2}\mathcal{M}'_{HF}(\Psi_{LF})(\Psi_{HF}-\Psi_{LF})$ can become feasible in certain scenarios; for example, if the high- and low-fidelity models only differ in the computational grid used, the difference $\Psi_{HF}-\Psi_{LF}$ can be approximated using interpolation or patch recovery methods~\cite{BecVex05}.
